### 2019-03-10 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/abstract.tex

\documentclass[12pt,a4paper]{article}

\title{\bf Human-Computer Mobile Distributed Computing for Time Series Forecasting}

\author{Rumen Ketipov, Georgi Kostadinov, Plamen Petrov, \\
Iliyan Zankinski, Todor Balabanov\textsuperscript{[0000-0003-3139-069X]} \\
\\
Institute of Information and Communication Technologies \\
Bulgarian Academy of Sciences \\
\\
acad. Georgi Bonchev Str, block 2, office 514 \\
1113 Sofia, Bulgaria \\
\\
todorb@iinf.bas.bg \\
http://www.iict.bas.bg/}

\date{}

\begin{document}

\maketitle

Financial time series forecasting is an attractive and intensively researched area. One of the most interesting research directions in this field is the usage of artificial neural networks combined with evolutionary algorithms. In some studies evolutionary algorithms are used for artificial neural network topology optimization, but in other studies evolutionary algorithms are used for artificial neural network weights optimization. When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm. Even when such hybrid training algorithm is used it is in the class of supervised training. Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm. Evolutionary algorithms have a common advantage compared to the exact numerical methods and it is the possibility for parallel calculations. When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing. It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be combined with human evaluation of the future values. In this research human-computer based mobile distributed computing solution is proposed. The users are using Android application, which trains artificial neural network with evolutionary algorithms and back-propagation, but users are capable to vote for increase or decrease of the future values. The most successful forecast votes of the users are used in artificial neural network training process.

\end{document}